\documentclass[svgnames]{amsart}
\usepackage[paperwidth=6in, paperheight=8in, top = 20mm, bottom = 18mm, left=10mm, right = 10mm]{geometry}

\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{graphicx, tikz}
\usepackage{mathtools, physics}

\usepackage{enumitem}
\setlist[enumerate,1]{label=\arabic*.}

\usepackage[default,bold]{sourceserifpro}
\usepackage[T1]{fontenc}
\usepackage[]{fouriernc}

\usepackage{xcolor}
\usepackage{scrlayer-scrpage}
\ohead{\color{blue!35!black} \scshape VM}
\cfoot*{\scriptsize\pagemark}

\renewcommand{\th}{\textsuperscript{th}}

\setlength{\parindent}{0pt}

\title[]{Probability and Statistics -- Problem Set 5}

\DeclareMathOperator{\Prob}{P}
\DeclareMathOperator{\EV}{\mathbb E}
\DeclareMathOperator{\Var}{\mathbb V}


\begin{document}
\maketitle
\begin{enumerate}[leftmargin=*, itemsep=0.3em]
\item Suppose that the number of items produced in a factory during one week is a random variable with mean $500$ and variance $100$.
\begin{enumerate}
	\item What can be said about the probability that this week's production will be at least $1000$?
	\item What can be said about the probability that this week's production will be between $400$ and $600$?
\end{enumerate}

\item If the time required to complete a task is a random variable with mean $20$ minutes and standard deviation $3$ minutes, find the smallest time frame such that the probability that the task will be completed within the time frame is at least $0.75$.

\item If $X \sim B(6, p)$, and $\Prob[X = 2] = 9 \Prob[X = 4]$, then find $\Prob[X \le 3]$.

\item An urn contains $4$ red, $4$ green, and $2$ blue marbles. One marble is drawn at random from the urn, observed, and placed back. Find the probability that in $8$ such draws
\begin{enumerate}
    \item exactly $3$ red marbles are drawn.
    \item not more than $3$ red marbles are drawn.
    \item at least $3$ red marbles are drawn.
    \item not all marbles drawn are red.
\end{enumerate}

\item The number of visitors to a webpage per minute follows a Poisson distribution. If the average number of visitors per minute is $4$, what is the probability that the webpage receives
\begin{enumerate}[label=(\roman*)]
	\item exactly four visitors in one minute?
	\item at least two visitors in one minute?
\end{enumerate}

\item If $X$ is a Poisson variate such that $3 \Prob[X = 2] = 2\Prob[X = 1]$, then what is $E[X]$?

\item  It is observed in a communication channel that $90\%$ of the messages sent are received without any error. Find the probability that among $18$ messages sent through the channel
\begin{enumerate}[label=(\roman*)]
	\item at least $16$ are received without any error
	\item at most $14$ are received without any error.
\end{enumerate}

\item In a certain factory producing blades, there is a small probability of $\frac{1}{500}$ for any blade to be defective. The blades are supplied in packets of $10$. Calculate the approximate number of packets containing
\begin{enumerate}[label=(\roman*)]
	\item no defective
	\item one defective
	\item two defective
	\item at least two defective
\end{enumerate}
blades in a consignment of $10000$ packets.

\item One per thousand of a population is subject to certain kinds of accidents each year. Given that an insurance company has insured $5,000$ persons from the population, find the probability that at most $2$ of them will incur this accident.

\item An airline company, having observed that $5\%$ of the persons making reservations on a flight do not show up for the flight, sells $100$ seats on a plane that has $95$ seats. What is the probability that there will be a seat available for every person who shows up for the flight?

\item If $X$ is a binomial variate with mean $3$, such that $\Prob[X = 3] = 2\Prob[X = 2]$, find $V[X]$.

\item $X$ is a Poisson variate and the probability that $X$ is even is twice the probability that it is odd. Determine $\Prob[X = 0]$.

\item  If $X$ is exponentially distributed with $E[X] = 2$, then find the value of $a$ such that $\Prob[X \le a] = \Prob[X \ge a]$.\\
{\small\color{blue!40!black}
		Note: This point $a$ is the \emph{median} of $X$.
}

\item Suppose that the duration in minutes of a phone call follows an exponential distribution with mean $5$ minutes.
\begin{enumerate}
\item  Find the probability that the duration of a particular call
\begin{enumerate}
	\item will exceed $5$ minutes
	\item will be between $5$ and $6$ minutes
	\item will be less than $3$ minutes
	\item will be less than $6$ minutes, given that it was greater than $3$ minutes.
\end{enumerate}
\item Suppose exactly $100$ such phone calls are received every day. Find the probability that on a given day
\begin{enumerate}
	\item every phone call lasted longer than $8$ minutes
	\item at least $4$ phone calls lasted longer than $8$ minutes.
\end{enumerate}
\item Find the probability that in one week, there are at least $3$ days on each which there were at least $4$ phone calls that lasted longer than $8$ minutes.
\end{enumerate}

\item A random variable $X$ follows an exponential distribution with parameter $\alpha = 3$. Compute
\begin{enumerate}
	\item $\Prob[2X > 1]$.
	\item $\Prob[X > s + t \mid X > t]$, where $s, t > 0$.
\end{enumerate}

\item The daily consumption of milk in a city, in excess of 20,000 gallons, is distributed as a Gamma variate with parameters $\alpha = \frac 1 {10000}$ and $r = 2$. The city has a daily stock of $30000$ gallons. What is the probability that the stock is insufficient on a particular day?

\item If $K \sim U[0, 5]$, then what is the probability that the roots of the equation $4x^2 + 4Kx + K + 2 = 0$ are real?

\item A book of $200$ pages contains $100$ misprints distributed randomly throughout its pages. If one page is selected at random and examined, what is the probability that it contains
\begin{enumerate}
	\item no misprints?
	\item at least $2$ misprints?
\end{enumerate}

\item Suppose that the chances of a traffic accident in a street of a city in a day is $0.001$. In how many days in a year can we expect
\begin{enumerate}
    \item no accidents?
    \item at least one accident, if there are $100$ such streets in a city?
\end{enumerate}
If there are $5$ such cities in a state, what is the probability that at least one city will have at least one accident in a day?

\item Buses start from a certain station at intervals of $45$ minutes. What is the probability that a person reaching the station at a random point in time will have to wait for at least $30$ minutes?

\item In a bombing action, there is a $50\%$ chance that a bomb will hit a target. Two direct hits are needed to destroy the target completely. Estimate the minimum number of bombs to be dropped to give $90\%$ or better chance of completely destroying the target.
\end{enumerate}
\end{document}