% !TEX program = xelatex

\documentclass[svgnames]{amsart}
\usepackage[paperwidth=6in, paperheight=8in, top = 20mm, bottom = 18mm, left=10mm, right = 10mm]{geometry}

\usepackage{amsmath, amsfonts, amssymb, amsthm}

\usepackage{enumitem}
\setlist[enumerate,1]{label=\arabic*.}

\usepackage[T1]{fontenc}
\usepackage[sfdefault, lining, scale=0.9]{FiraSans}
\usepackage{kmath}

\usepackage{xcolor}
\usepackage{scrlayer-scrpage}
\ohead{\color{blue!35!black} \scshape VM}
\cfoot*{\pagemark}

\renewcommand{\th}{\textsuperscript{th}}

\setlength{\parindent}{0pt}
\linespread{1.2}

\title{MAT 2155: Problem Set 4}
\date{}

\begin{document}
\maketitle
\begin{enumerate}[leftmargin=*]
\item The coefficient of $x^6$ in the expansion of
\begin{enumerate}[label=(\roman*)]
	\item $(1 + x)^8$.
	\item $(1 - x)^{-8}$.
	\item $(1 + x^2)^4$.
	\item $(1 - x^2)^{-4}$.
	\item $(1 + x + x^2)^4$.
	\item $(1 + x^2)(1 + x^2 + x^4 + \cdots)$.
\end{enumerate}

\item Using generating functions, find the number of ways of selecting $6$ objects from
\begin{enumerate}[label=(\roman*)]
	\item $8$ distinct objects.
	\item $8$ types of object, with any number of objects of each type.
	\item $4$ types of objects, such that zero or two objects are chosen from each type.
	\item $4$ types of objects, such that an even number of objects is chosen from each type.
	\item $4$ types of objects, such that not more than two objects are chosen from each type.
	\item $2$ types of objects, such that zero or two objects of the first type are chosen, and an even number of objects is chosen from the second type.
\end{enumerate}

\item Using generating functions, find the number of integer solutions of the equation
\begin{enumerate}[label=(\roman*)]
	\item $x_1 + \cdots + x_8 = 6$, $x_i \in \{0, 1\}$, $i = 1, \ldots, 8$.
	\item $x_1 + \cdots + x_8 = 6$, $x_i \ge 0$, $i = 1, \ldots, 8$.
	\item $x_1 + \cdots + x_4 = 6$, $x_i \in \{0, 2\}$, $i = 1, \ldots, 4$.
	\item $x_1 + \cdots + x_4 = 6$, $x_i \ge 0$, $x_i$ is even, $i = 1, \ldots, 4$.
	\item $x_1 + \cdots + x_4 = 6$, $0 \le x_i \le 2$, $i = 1, \ldots, 4$.
	\item $x + y = 6$, $x \in \{0, 2\}$, $y \ge 0$, $y$ is even.
\end{enumerate}

\item Number of ways of distributing $30$ identical objects into $3$ distinct boxes such that no box is empty.

\item Number of ways of distributing $30$ identical marbles into $6$ boxes with at most $10$ marbles in the first box.

\item Number of ways of selecting $12$ flowers for a bouquet from roses, lilacs, tulips, and lilies, with between $2$ and $5$ of each kind.

\item Number of ways to select $10$ marbles from a large pile of red, white, and blue marbles if
\begin{enumerate}[label=(\roman*)]
	\item the selection has at least $2$ marbles of each colour.
	\item the selection has at most $2$ red marbles.
	\item the selection has an even number of blue marbles.
\end{enumerate}

\item Number of ways to place an order of $12$ chocolate sundaes if there are $5$ types of sundaes, and at most $4$ sundaes of one type are allowed.

\item Number of ways to get a sum of $25$ when $10$ distinct dice are rolled.

\item Number of ways to select $300$ chocolate candies from $7$ types of candy if each type comes in boxes of $20$, and at least $1$ but not more than $5$ boxes of each type are chosen.

\item Number of ways of distributing $30$ distinct objects into $3$ boxes such that no box is empty.

\item Using generating functions, find the number of $r$-permutations of objects chosen from unlimited supplies of $n$ types of objects.

\item Number of $r$-digit quaternary sequences (with digits $0$, $1$, $2$, $3$) having an even number of $0$s and an odd number of $1$s.

\item Write the exponential generating function for the number of arrangements of $k$ objects chosen from $n$ types with at most $4$ objects of each type.

\item Number of $n$-digit ternary sequences with
\begin{enumerate}[label=(\roman*)]
	\item an even number of $0$s.
	\item an even number of $0$s and an even number of $1$s.
	\item $0$ and $1$ occurring a positive even number of times.
	\item at least one $0$ and at least one $1$.
	\item the total number of $0$s and $1$s being even.
	\item no digit occurs exactly twice.
\end{enumerate}
\end{enumerate}
\end{document}