% !TEX program = xelatex

\documentclass[svgnames]{amsart}
\usepackage[paperwidth=6in, paperheight=8in, top = 20mm, bottom = 18mm, left=10mm, right = 10mm]{geometry}

\usepackage{amsmath, amsfonts, amssymb, amsthm}

\usepackage{enumitem}
\setlist[enumerate,1]{label=\arabic*.}

\usepackage[T1]{fontenc}
\usepackage[sfdefault, lining, scale=0.9]{FiraSans}
\usepackage{kmath}

\usepackage{xcolor}
\usepackage{scrlayer-scrpage}
\ohead{\color{blue!35!black} \scshape VM}
\cfoot*{\pagemark}

\renewcommand{\th}{\textsuperscript{th}}

\setlength{\parindent}{0pt}
\linespread{1.2}

\title{MAT 2155: Problem Set 2}
\date{}

\begin{document}
\maketitle
\begin{enumerate}[leftmargin=*]
\item Number of rearrangements of the word \texttt{SESQUIPEDALIAN}.

\item Number of anagrams of the word \texttt{INDEPENDENCE} which
\begin{enumerate}[label=(\roman*)]
	\item start with the letter \texttt{C}.
	\item all the vowels occur together.
	\item no two vowels occur together.
	\item begin in \texttt{I} and end in \texttt{E}.
\end{enumerate}

\item Number of positive integers less than $1000$ that have distinct digits.

\item Number of ways of choosing $4$ cards from a pack of $52$ playing cards
\begin{enumerate}[label=(\roman*)]
	\item with no restrictions.
	\item with all four cards from the same suit.
	\item with all four cards from different suits.
	\item such that all four cards are face cards.
	\item such that $2$ are red and $2$ are black.
	\item such that all four cards are of the same colour.
\end{enumerate}

\item There are $9$ different books on a bookshelf, $4$ red and $5$ green. In how many different ways can the books be arranged if
\begin{enumerate}[label=(\roman*)]
	\item there are no restrictions?
	\item the red books must be together, and the green books must be together?
	\item the red books must be together?
	\item no two books of the same colour must be together?
\end{enumerate}

\item A shop sells six different flavours of ice cream. In how many ways can a customer choose $4$ ice cream cones
\begin{enumerate}[label=(\roman*)]
	\item all of different flavours?
	\item not necessarily of different flavours?
	\item of $2$ or $3$ different flavours?
	\item exactly $3$ different flavours?
\end{enumerate}

\item Number of ways of buying a total of $7$ fruits from a shop selling apples, oranges, and strawberries, if you must buy at least one fruit of each type.

\item Number of ways of selecting $10$ marbles from a pile of red, blue, and green marbles if there must be
\begin{enumerate}[label=(\roman*)]
	\item at least $5$ red marbles.
	\item at most $5$ red marbles.
\end{enumerate}

\item Number of integer solutions of the equation $x_1 + x_2 + x_3 + x_4 = 18$ with
\begin{enumerate}[label=(\roman*)]
	\item $x_i \ge 0$, $i = 1, 2, 3, 4$.
	\item $x_i \ge 1$, $i = 1, 2, 3, 4$.
	\item $x_i \ge i$, $i = 1, 2, 3, 4$.
\end{enumerate}

\item Number of binary sequences of length $10$ consisting of a run of $1$s followed by a run of $0$s followed by a run of $1$s followed by another run of $0$s.

\item Number of ways to distribute $15$ identical objects into $4$ different boxes such that the number of objects in the fourth box is a multiple of $3$.
\end{enumerate}
\end{document}